---
title: "scrape"
author: "jjtimmons"
date: "7/21/2018"
---

# NFL Player Data Scrape

This is a script for scraping NFL player data from [www.pro-football-reference.com](www.pro-football-reference.com).

## Scrape

Will proceed from 2010 through 2017, collecting NFL player data and fantasy points accumulated from each year. Example of the html for a single `tr` on the site is below:

```{html}
<tr data-row="113">
  <th scope="row" class="right " data-stat="ranker" csk="111">1</th>
  <td class="left " data-append-csv="AddaJo00" data-stat="player" csk="Addai,Joseph"><a href="/players/A/AddaJo00.htm">Joseph Addai</a></td>
  <td class="left " data-stat="team"><a href="/teams/clt/2008.htm" title="Indianapolis Colts">IND</a></td>
  <td class="right " data-stat="fantasy_pos" csk="20">RB</td>
  <td class="right " data-stat="age">25</td>
  <td class="right " data-stat="g">12</td>
  <td class="right " data-stat="gs">12</td>
  <td class="right " data-stat="pass_cmp">0</td>
  <td class="right " data-stat="pass_att">0</td>
  <td class="right " data-stat="pass_yds">0</td>
  <td class="right " data-stat="pass_td">0</td>
  <td class="right " data-stat="pass_int">0</td>
  <td class="right " data-stat="rush_att">155</td>
  <td class="right " data-stat="rush_yds">544</td>
  <td class="right " data-stat="rush_yds_per_att">3.51</td>
  <td class="right " data-stat="rush_td">5</td>
  <td class="right " data-stat="targets">39</td>
  <td class="right " data-stat="rec">25</td>
  <td class="right " data-stat="rec_yds">206</td>
  <td class="right non_qual" data-stat="rec_yds_per_rec">8.24</td>
  <td class="right " data-stat="rec_td">2</td>
  <td class="right " data-stat="fantasy_points">117</td>
  <td class="right " data-stat="fantasy_points_ppr"></td>
  <td class="right " data-stat="draftkings_points">147.5</td>
  <td class="right " data-stat="fanduel_points">128.5</td>
  <td class="right " data-stat="vbd"></td>
  <td class="right " data-stat="fantasy_rank_pos">39</td>
  <td class="right " data-stat="fantasy_rank_overall"></td>
</tr>
```

I'm going to preserve/re-use the data headers to keep it simple. Everything that's a data-stat attribute will become a column header in the R columns. Create a vector of years, scrape each from pro-football-reference and save to a CSV.

```{r}
library(rvest)
library(stringr)
library(xts)

setwd("~/Documents/GitHub/ff/data")

FIRST_YEAR <- 1997
LAST_YEAR <- 2017

clean_name <- function(row) {
  row[[2]] <- str_replace_all(row[[2]], "[*+]", "")
  row
}

scrape_year <- function(year) {
  url <- paste0("https://www.pro-football-reference.com/years/", year, "/fantasy.htm")
  page <- read_html(url)
  
  player_table <- page %>%
    html_nodes("#fantasy") %>%
    html_table()
  player_table <- player_table[[1]] # only one per page
  
  headers <- player_table[1,] # first row is headers
  headers[2] <- "Name" # idk why but this is blank by default
  colnames(player_table) <- headers # set headers
  player_table <- player_table[nchar(player_table$Name) > 1, ] # remove title rows
  
  player_table <- apply(player_table, 1, clean_name) # get rid of extraneous symbols
  player_table <- t(player_table)
  player_table <- data.frame(player_table)
  
  player_table$Year <- rep(year, nrow(player_table)) # set year in the Year column
  player_table
}

player.data <- lapply(seq(from=FIRST_YEAR, to=LAST_YEAR), scrape_year)
player.data <- do.call(rbind, player.data) # https://www.r-bloggers.com/concatenating-a-list-of-data-frames/
save(player.data, file = paste0("player_data.raw.Rda"))
rm(list=ls())
```

## Clean

Clean the data by converting the column types, moving Year to the second column (needed by plm). Add an "xp" field based on the number of years that the player has been in the NFL.

```{r}
load("~/Documents/GitHub/ff/data/player_data.raw.Rda")

# convert all the columns to numeric that should be
numerics <- c("Year", "FantPt", "Yds", "Yds.1", "Yds.2", "TD", "TD.1", "TD.2", "Int", "Rk", "G", "GS", "Cmp", "Att", "FantPt", "PosRank", "Rec", "Tgt", "Y.R",  "PPR", "DKPt", "Age", "FDPt")
for (n in numerics)
  player.data[,n] <- as.numeric(as.character(player.data[,n]))

# sort by name then year
player.data <- player.data[order(player.data$Name, player.data$Year),]

# all lowercase
names(player.data) <- tolower(names(player.data))

# team as a factor
player.data$tm <- as.factor(player.data$tm)

# reorder the columns
first.cols <- c("name", "year", "fantpos", "age")
ordered.cols <- c(first.cols, setdiff(tolower(numerics), first.cols))
player.data <- player.data[, ordered.cols]

# save to file system
save(player.data, file = paste0("player_data.Rda"))
rm(list=ls())
```

